---
title: "R Notebook"
output: html_notebook
---

These are the libraries you will need. RStudio should prompt you to install them
when it opens, but if not you can type install.packages("readxl") foe each of 
them in the console
```{r}
library(readxl)
library(ggplot2)
library(stringr)
library(dplyr)
library(RColorBrewer)
library(cowplot)
library(ggpubr)
```

#Make sure the xlsx file is in the same folder as this R-file
```{r}
# lines of code that start with a # are comments. That means they are notes that
# wont actually run. It's to help you remember what each line of code does

# This reads in the data
BMDData <- read_excel("BMDEstimation.xlsx", skip = 1)

# This removes rows with missing data
BMDData <- na.omit(BMDData)

# This breaks your Vert column up by Position and number
verts = as.data.frame(str_split_fixed(BMDData$Vert, ' ', 2))
BMDData$VertPos = verts$V1
BMDData$VertNum = verts$V2

# This puts the data in "front-middle-back" order instead of alphabetical order
BMDData <- mutate(BMDData, VertPos = factor(VertPos, levels=c("Front", "Middle", "Back")))

BMDData <- mutate(BMDData, Vert = factor(Vert, levels=c("Front 1","Front 2","Front 3","Front 4","Front 5", "Middle 1","Middle 2","Middle 3","Middle 4","Middle 5","Back 1","Back 2","Back 3","Back 4","Back 5")))
```

#Plotting
We are using the package ggplot for plotting. It requires two basic lines to
start. The ggplot line, and the geometry line (marked below). The ggplot
line tells the functuon which data your want to plot. Your data frame, x and y
variables, groupings, etc. The geom tells it what tupe of plot you want (boxplot,
scatter plot, violin plot, etc.)

#Boxplots (Looking at position)
```{r}
ggplot(data = BMDData, aes(x = VertPos, y = BMD))+ #ggplot line
  geom_boxplot() #geometry line
  
# Boxplot by position
my_comparisons <- list( c("Front", "Middle"), c("Middle", "Back"), c("Front", "Back") )

bmdbox <- ggplot(data = BMDData, aes(x = VertPos, y = BMD, fill = VertPos))+
  geom_boxplot()+ 
  scale_fill_manual(values = c("#E69F00", "#56B4E9", "#9955BB"))+ # this adds a color palette
  scale_x_discrete(labels = c("Front" = "Anterior", "Middle", "Back"= "Posterior"))+ # this sets your categorical labels on the x axis
  xlab("Position")+ # this sets the general x axis label
  ylab("% Hydroxyapatite")+ # this sets the y axis label
  ylim(22,72)+ # this sets the limits of the y axis
  theme_classic()+ # classic theme
  theme(legend.position="none")+ 
  stat_compare_means(comparisons = my_comparisons, label.y = c(58, 62, 66), color = "black", size = 2.5) # statistics to compare means
show(bmdbox)
ggsave("BMDBox.png", height = 4, width = 4, units = "in") # saves png to your computer in the same folder as this R script


#Boxplot by position and number
ggplot(data = BMDData, aes(x = VertPos, y = BMD, fill = VertNum))+
  geom_boxplot()+
  theme_classic()+
  scale_fill_brewer(palette = "GnBu")

```
R Colors: https://sape.inf.usi.ch/quick-reference/ggplot2/colour
R Color Palettes: https://r-graph-gallery.com/38-rcolorbrewers-palettes.html

```{r}
# This creates a linear equation for your data 
lm_eqn <- function(y,x){
    m <- lm(y ~ x);
    eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2, 
         list(a = format(unname(coef(m)[1]), digits = 2),
              b = format(unname(coef(m)[2]), digits = 2),
             r2 = format(summary(m)$r.squared, digits = 3)))
    as.character(as.expression(eq));
}
```


#Scatterplots (adding body length)
```{r}
# This creates data subsets by position
Front <- filter(BMDData, VertPos == "Front")
Middle <- filter(BMDData, VertPos == "Middle")
Back <- filter(BMDData, VertPos == "Back")

frontModel = lm(BMD ~ Length, data = Front)
anova(frontModel)

middleModel = lm(BMD ~ Length, data = Middle)
anova(middleModel)

backModel = lm(BMD ~ Length, data = Back)
anova(backModel)

print(lm_eqn(Front$BMD, Front$Length))

# this creates 3 scatter plots (p1, p2, p3) one for each position
p1 <- ggplot(data = Front, aes(x = Length, y = BMD))+
  geom_point(color = "#E69F00", shape=1, size = 2)+ # this sets the color, shape, and size of points
  ylim(min(BMDData$BMD), max(BMDData$BMD)+2)+ # this sets the upper and lower limit of the y-axis
  xlab("")+
  ylab("")+ # this labels your y-axis
  geom_smooth(method = "lm", se = FALSE, color = "black")+ #this adds a trendline
  theme_classic()+ #this adds the classic theme
  geom_text(size = 3.5, x = 75, y = 72, label = lm_eqn(Front$BMD, Front$Length), parse = TRUE, color = "black") # this adds your equation to the plot 
show(p1)

p2 <- ggplot(data = Middle, aes(x = Length, y = BMD))+
  geom_point(color = "#56B4E9", shape=1, size = 2)+
  theme_classic()+
  ylim(min(BMDData$BMD), max(BMDData$BMD)+2)+
  xlab("")+
  ylab("")+
  geom_smooth(method = "lm", se = FALSE, color = "black")+
  geom_text(size = 3.5, x = 75, y = 72, label = lm_eqn(Middle$BMD, Middle$Length), parse = TRUE, color = "black")
  

p3 <- ggplot(data = Back, aes(x = Length, y = BMD))+
  geom_point(color = "#9955BB", shape=1, size = 2)+
  theme_classic()+
  ylim(min(BMDData$BMD), max(BMDData$BMD)+2)+
  xlab("")+
  ylab("")+
  geom_smooth(method = "lm", se = FALSE, color = "black")+
  geom_text(size= 3.5, x = 75, y = 72, label = lm_eqn(Back$BMD, Back$Length), parse = TRUE, color = "black")
 

# the plot_grid function allows you to put a bunch of plots together
plot_grid(bmdbox,p1,p2,p3, nrow = 4)
ggsave("BMDScatterBox.png", height = 12, width = 3, units = "in")

```

Very simple statistics
```{r}
# Does position and length predict BMD?
boneModel <- lm(BMD ~ VertPos + Length, data = BMDData)
anova(boneModel)

#Which positions are different from each-other
lengthModel1 <- pairwise.t.test(BMDData$BMD, BMDData$VertPos, p.adj = "bonf")
lengthModel1

boneModelAnt <- lm(BMD ~ Length, data = Front)
anova(boneModelAnt)

boneModelMid <- lm(BMD ~ Length, data = Middle)
anova(boneModelMid)

boneModelPos <- lm(BMD ~ Length, data = Back)
anova(boneModelPos)
```

