---
title: "R Notebook"
output: html_notebook
---

These are the libraries you will need. RStudio should prompt you to install them
when it opens, but if not you can type install.packages("readxl") foe each of 
them in the console
```{r}
library(readxl)
library(ggplot2)
library(paletteer)
library(stringr)
library(dplyr)
library(RColorBrewer)
library(cowplot)
library(ggpubr)
```

Black Background (run this chunk to save theme to your environment)
```{r}
theme_black = function(base_size = 12, base_family = "") {
 
  theme_grey(base_size = base_size, base_family = base_family) %+replace%
 
    theme(
      # Specify axis options
      axis.line = element_blank(),  
      axis.text.x = element_text(size = base_size*0.8, color = "white", lineheight = 0.9),  
      axis.text.y = element_text(size = base_size*0.8, color = "white", lineheight = 0.9),  
      axis.ticks = element_line(color = "white", size  =  0.2),  
      axis.title.x = element_text(size = base_size, color = "white", margin = margin(0, 10, 0, 0)),  
      axis.title.y = element_text(size = base_size, color = "white", angle = 90, margin = margin(0, 10, 0, 0)),  
      axis.ticks.length = unit(0.3, "lines"),   
      # Specify legend options
      legend.background = element_rect(color = NA, fill = "black"),  
      legend.key = element_rect(color = "black",  fill = "black"),  
      legend.key.size = unit(1.2, "lines"),  
      legend.key.height = NULL,  
      legend.key.width = NULL,      
      legend.text = element_text(size = base_size*0.8, color = "white"),  
      legend.title = element_text(size = base_size*0.8, face = "bold", hjust = 0, color = "white"),  
      legend.position = "right",  
      legend.text.align = NULL,  
      legend.title.align = NULL,  
      legend.direction = "vertical",  
      legend.box = NULL, 
      # Specify panel options
      panel.background = element_rect(fill = "black", color  =  NA),  
      panel.border = element_rect(fill = NA, color = "white"),  
      panel.grid.major = element_blank(),  
      panel.grid.minor = element_blank(),  
      panel.margin = unit(0.5, "lines"),   
      # Specify facetting options
      strip.background = element_rect(fill = "grey30", color = "grey10"),  
      strip.text.x = element_text(size = base_size*0.8, color = "white"),  
      strip.text.y = element_text(size = base_size*0.8, color = "white",angle = -90),  
      # Specify plot options
      plot.background = element_rect(color = "black", fill = "black"),  
      plot.title = element_text(size = base_size*1.2, color = "white"),  
      plot.margin = unit(rep(1, 4), "lines")
 
    )
 
}
```

#Make sure the xlsx file is in the same folder as this R-file
```{r}
# lines of code that start with a # are comments. That means they are notes that
# wont actually run. It's to help you remember what each line of code does

# This reads in the data
library(readxl)
DLTdv8_Data <- read_excel("DLTdv8 Model Videos Data.xlsx")
View(DLTdv8_Data)


# This removes rows with missing data
DLTdv8_Data <- na.omit(DLTdv8_Data)

```

#Plotting
We are using the package ggplot for plotting. It requires two basic lines to
start. The ggplot line, and the geometry line (marked below). The ggplot
line tells the function which data you want to plot. Your data frame, x and y
variables, groupings, etc. The geom tells it what tupe of plot you want (boxplot,
scatter plot, violin plot, etc.)

#Boxplots (Looking at position)
```{r}
# Frequency and Tail Amplitude
ggplot(data = DLTdv8_Data, aes(x = Freq.Hz, y = Y1amp, color = Model))+ #ggplot line
  geom_point(size = 3)+
  theme_classic()+
  scale_color_manual(values = c("#332288", "#117733", "#44AA99", "#88CCEE",
                               "#DDCC77", "#CC6677", "#AA4499", "#882255"))+
  xlab("Actuator Frequency")+
  ylab("Tail Amplitude")+
  geom_smooth(method="lm", se=FALSE) #geometry line
ggsave("FreqvTailAmp.png", height = 11.55, width = 13.71, units = "in")

# Frequency at Tail vs Head
ggplot(data = DLTdv8_Data, aes(x = Y2freq, y = Y1freq, color = Model))+ #ggplot line
  geom_point(size = 3)+
  theme_classic()+
  scale_color_manual(values = c("#332288", "#117733", "#44AA99", "#88CCEE",
                               "#DDCC77", "#CC6677", "#AA4499", "#882255"))+
  xlab("Head Frequency")+
  ylab("Tail Frequency")+
  geom_smooth(method="lm", se=FALSE) #geometry line
ggsave("HeadvTailFreq.png", height = 11.55, width = 13.71, units = "in")

# Frequency and Wavespeed
ggplot(data = DLTdv8_Data, aes(x = Freq.Hz, y = waveSpeed, color = Model)) + #ggplot line
  geom_point(size = 5) +
  theme_classic()+
  scale_color_manual(values = c("#332288", "#117733", "#44AA99", "#88CCEE",
                               "#DDCC77", "#CC6677", "#AA4499", "#882255"))+
  geom_line() #geometry line
ggsave("WaveSpeedvFreq.png", height = 10, width = 15, units = "in")

# Amplitude Ratio
ggplot(data = DLTdv8_Data, aes(x = Y2amp, y = Y1amp, color = Model))+ #ggplot line
  geom_point(size = 3)+
  theme_classic()+
  scale_color_manual(values = c("#332288", "#117733", "#44AA99", "#88CCEE",
                               "#DDCC77", "#CC6677", "#AA4499", "#882255"))+
  xlab("Head Amplitude")+
  ylab("Tail Amplitude")+
  geom_smooth(method="lm", se=FALSE) #geometry line
ggsave("HeadTailAmplitudes.png", height = 11.55, width = 13.71, units = "in")


```
R Colors: https://sape.inf.usi.ch/quick-reference/ggplot2/colour
R Color Palettes: https://r-graph-gallery.com/38-rcolorbrewers-palettes.html

```{r}
lm_eqn <- function(y,x){
    m <- lm(y ~ x);
    eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2, 
         list(a = format(unname(coef(m)[1]), digits = 2),
              b = format(unname(coef(m)[2]), digits = 2),
             r2 = format(summary(m)$r.squared, digits = 3)))
    as.character(as.expression(eq));
}
```


#Scatterplots ()
```{r}
# Separating by Density Distributions
library(dplyr)
AModels <- dplyr::filter(DLTdv8_Data, Model == "A")
BModels <- dplyr::filter(DLTdv8_Data, Model == "B")
CModels <- dplyr::filter(DLTdv8_Data, Model == "C")
DModels <- dplyr::filter(DLTdv8_Data, Model == "D")
EModels <- dplyr::filter(DLTdv8_Data, Model == "E")
FModels <- dplyr::filter(DLTdv8_Data, Model == "F")
GModels <- dplyr::filter(DLTdv8_Data, Model == "G")
HModels <- dplyr::filter(DLTdv8_Data, Model == "H")

# Amplitude Ratios
amplitude_ratios <- DLTdv8_Data %>%
  group_by(Model, Freq.Hz) %>%
  summarise(
    amp_ratio = Y1amp / Y2amp,
    .groups = "drop"
  )

ggplot(amplitude_ratios,
       aes(x = Freq.Hz, y = amp_ratio, color = Model)) +
  geom_point(size = 3) +
  geom_line() +
  labs(x = "Frequency (Hz)",
       y = "Amplitude Ratio (Tail / Head)") +
  theme_black()
ggsave("AmpRatiosvFreq.png", height = 5 , width = 10, units = "in")

amplitude_ratios <- amplitude_ratios %>%
  mutate(
    model = factor(Model),
    frequency = factor(Freq.Hz)
  )

amplitude_ratios$model <- factor(amplitude_ratios$model)


# Do model and frequency predict amplitude ratios?
lm1 <- lm(amp_ratio ~ model + Freq.Hz, data = amplitude_ratios)
anova(lm1)

# model-specific frequency response
lm2 <- lm(amp_ratio ~ model * Freq.Hz, data = amplitude_ratios)
anova(lm2)

anova_model <- aov(amp_ratio ~ model, data = amplitude_ratios)
summary(anova_model)

tukey_models <- TukeyHSD(anova_model, "model")

print(tukey_models)

```

